\section*{Exam 1 Review}
\addcontentsline{toc}{section}{Exam 1 Review}
\setcounter{counter}{0}
\begin{revex}{Vector Operations}
\begin{enumerate}
\item Find the cross product: $\bmat{3\\2\\5}\times\bmat{1\\2\\-2}.$
\vspace{1em}
\item Find the cross product: $\bmat{-1\\2\\-3}\times \bmat{-4\\-2\\1}. $
\vspace{1em}
\item Let $\vcv=\bmat{1\\2\\3} $ and $\vcw=\bmat{-1\\-1\\4\\}$. Find $\proj{\vcv}{\vcw}$.
\end{enumerate}
\end{revex}

\begin{revex}{Constructing Vectors}
For each of the properties below, construct a vector $\vec{v}\in\mathbb{R}^3$ with the given properties, or explain why no such vector exists.

\begin{enumerate}
    \item $|| \vec{v} || = 1 $ and $\vec{v}$ has a projection of zero onto each of $\vec{i}, \vec{j},$ and $\vec{k}$
    \vspace{1em}
        \item $|| \vec{v} || = 1 $ and $\vec{v}$ has a projection of zero onto exactly two of $\vec{i}, \vec{j},$ or $\vec{k}$
    \vspace{1em}
    \item $|| \vec{v} || = 1 $ and $\vec{v}$ has a projection of zero onto exactly one of $\vec{i}, \vec{j},$ or $\vec{k}$
    \vspace{1em}
    \item $|| \vec{v} || = 1 $ and $\vec{v}$ has a projection of zero onto none of $\vec{i}, \vec{j},$ or $\vec{k}$
    
\end{enumerate}
\end{revex}

\begin{revex}{Equation of a Plane}
Find the equation in standard form of a plane that contains the point $(1,2,3)$ and is parallel to the vectors $$\vcd_1=\bmat{-1\\2\\1}\text{ and }\vcd_2=\bmat{-3\\2\\-4}$$
\end{revex}

\begin{revex}{Equation of a Line}
Consider the two planes $x+2y-z=4$ and $3x-y=1$. Find the equation of the line that is the intersection of these two planes in parametric form.
\end{revex}

\begin{revex}{Point to Plane}
Find the distance between the point $(-1,-1,-1)$ and the plane $x+y+z=3$.
\end{revex}

\begin{revex}{Point to Line}
Find the distance between the point $(-1,1,-1)$ and the line $L(t)=\bmat{t\\-t\\t+1}$.
\end{revex}

\begin{revex}{Line to Line 1}
Consider the lines $$L_1(t)=\bmat{t+2\\-t+3\\2t}\text{ and }L_2(t)=\bmat{t-1\\-t-1\\2t+4}. $$
\begin{enumerate}
\item Are $L_1(t)$ and $L_2(t)$ parallel, skew, or do they intersect?
\vspace{1em}
\item Give the distance between $L_1(t)$ and $L_2(t)$.
\end{enumerate}
\end{revex}

\begin{revex}{Line to Line 2}
Consider the lines $$L_1(t)=\bmat{3t\\2t+1\\t}\text{ and }L_2(t)=\bmat{-t\\3t\\-t+4}. $$
\begin{enumerate}
\item Are $L_1(t)$ and $L_2(t)$ parallel, skew, or do they intersect?
\vspace{1em}
\item Give the distance between $L_1(t)$ and $L_2(t)$.
\end{enumerate}
\end{revex}

\begin{revex}{PVA and Curvature}
Suppose a projectile moves through 3-dimensional space with velocity $$\vcv(t)=\bmat{3t^2\\t\\2}. $$
At time $t=0$, the projectile is at the origin.
\vspace{1em}
\begin{enumerate}
\item Find $\vcr(t)$, the position function for the projectile.
\vspace{1em}
\item Find $\vca(t)$, the acceleration function for the projectile.
\vspace{1em}
\item Find $\kappa(t)$, the curvature of the projectile.
\end{enumerate}
\end{revex}

\begin{revex}{Tangent Line}
Consider the function $$\vcr(t)=\bmat{t^2\\t\\-3t}. $$
Find the equation in parametric form of the line tangent to $\vcr(t)$ at $t=2$.
\end{revex}

\begin{revex}{Evolute}
Consider the function $$\vcr(t)=\bmat{2\cos(2t)\\2\sin(2t)\\3t}. $$
\begin{enumerate}
\item Find $\vcT(t)$, the tangent vector.
\vspace{1em}
\item Find $\vcN(t)$, the normal vector.
\vspace{1em}
\item Find $\kappa(t)$, the curvature.
\vspace{1em}
\item Find $\vcE(t)$, the evolute.
\end{enumerate}
\end{revex}

\begin{revex}{Arc Length}
Find the length of the function $$\vcr(t)=\bmat{\sqrt{t^3}\\t\\2t} $$ between $t=0$ and $t=1$.
\end{revex}