\section*{Exam 2 Review}
\addcontentsline{toc}{section}{Exam 2 Review}
\setcounter{counter}{0}
\begin{revex}{Partial Derivatives}
Let $f(x,y,z)=\frac{xyz}{\sin(x)}\ln(y)$.
\vspace{1em}
\begin{enumerate}
\item Find $f_x(x,y,z)$.
\vspace{1em}
\item Find $f_y(x,y,z)$.
\vspace{1em}
\item Find $f_z(x,y,z)$.
\vspace{1em}
\item Find $f_{zz}(x,y,z)$.
\end{enumerate}
\end{revex}

\begin{revex}{Tangent Plane}
Let $f(x,y)=x^2-y^2$. Find the equation of the plane tangent to $f$ at $(1,2)$.
\end{revex}

\begin{revex}{Critical Points}
Let $f(x,y)=x^2y-3xy+2y$. Find all of the critical points of $f$ and categorize them as maximums, minimums or saddle points according to the second derivative test.
\end{revex}

\begin{revex}{Lagrange Multipliers}
Suppose you are running a business that sells 3 products, product $X$, $Y$ and $Z$. Let $x,y,z$ be the quantity of each respective product. The total revenue your company makes is $$R(x,y,z)=-0.1x^2+30x-0.2y^2+45y-0.5z^2+60z.$$ Your company has a budget of $\$1,000$ for production. Products $X$ and $Z$ cost $\$5$ per unit and product $Y$ costs $\$10$ per unit. Use Lagrange Multipliers to find the optimal quantities of each product to maximize your revenue.
\end{revex}

\begin{revex}{Double Integral}
Let $R$ be the region bounded by $y=\cos(x)$, $y=\sin(x)$ and $x=0$. Evaluate $$\iint_R x \ dA. $$
\end{revex}

\begin{revex}{Triple Integral}
Let $R$ be the rectangle $[0,1]\times[0,2]\times[0,3].$ Evaluate $$\iiint_R xyz\ dV $$
\end{revex}

\begin{revex}{Change of Variables 1}
Let $R$ be the triangle with corners at $(0,0)$, $(-1,3)$ and $(1,2)$. Use the change of variables $x=u-2v$, $y=2u+v$ to evaluate $$\iint_R xy\ dA. $$
\end{revex}

\begin{revex}{Change of Variables 2}
Let $R$ be the upper half of the sphere with radius $4$. Use a change of variables to spherical coordinates to evaluate $$\iiint_R 3\ dV. $$
\end{revex}

\begin{revex}{Surface Area}
The surface $f(x,y)=1-1\sqrt{x^2+y^2}$ above the disk of radius $1$ in the $xy$-plane gives a cone with a height of 1 and a radius of 1. Verify that the surface area of this cone is in fact, $\sqrt{2}\pi$. Note that the surface area of the conical section of a cone with height $h$ and radius $r$ is $\pi r\sqrt{h^2+r^2}$
\end{revex}

\begin{revex}{Center of Mass}
Consider the region between the cone $z=2-2\sqrt{x^2+y^2}$ and the $xy$-plane. If the density of this region is $\rho(x,y,z)=z$, find the center of mass of the region. Hint: You probably want to use cylindrical coordinates to compute this. Also $\overline{x}$ and $\overline{y}$ can be computed by symmetry.
\end{revex}