\renewcommand\thesubsection{\thesection.\Alph{subsection}}
\setcounter{subsection}{17}
\subsection{Vector Operations and Properties Homework and Miscellaneous Practice}
\begin{exercise}{Definition of Unit Vector}
How can you tell if a vector is a unit vector? 
\end{exercise}

\begin{exercise}{Cross Product in 2-D}
The cross product is not defined in $2$-dimensional space. How do we take a cross product of two $2$-dimensional vectors?
\end{exercise}

\begin{exercise}{Flatland}
Suppose that you're a denizen of ``Flatland", a 2-dimensional space (i.e. a euclidean plane). Imagine that one day, a sphere passes through your plane. (i.e. imagine you can only percieve the $xy$-plane in $\bbr^3$ and a sphere moves from above the $xy$-plane to below the $xy$-plane). How would you perceive the sphere? (This question will \textit{not} be on the quiz).
\end{exercise}

\begin{exercise}{Operation Practice}
Let $\vcv=(1,2,3)$, $\vcu=(1,-1,3)$ and $\vcw=(-2,4,1)$.
\vspace{1em}
\begin{enumerate}
\item Find $||\vcv||$.
\vspace{1em}
\item Find $||\vcu||$.
\vspace{1em}
\item Find $||\vcw||$.
\vspace{1em}
\item Find $2\vcv-\vcu$.
\vspace{1em}
\item Find $\vcv\bullet\vcw$.
\vspace{1em}
\item Find $\vcv\times\vcu$.
\vspace{1em}
\item Find $\vcv_1$ such that $\vcv_1$ is a unit vector parallel to $\vcv$.
\vspace{1em}
\item Find the angle between $\vcw$ and $\vcv$ using the dot product.
\vspace{1em}
% \item Find the angle between $\vcu$ and $\vcv$ using the cross product.
% \vspace{1em}
\item Find $\proj{\vcv}{\vcw}$.
\vspace{1em}
\item Find $\proj{\vcw}{\vcv}$.
\end{enumerate}
\end{exercise}

\begin{exercise}{Magnitude and Dot Product}
Let $\vcv=(v_1,v_2,\ldots,v_n)$ be a vector in $\bbr_n$. Show that $\vcv\bullet\vcv=||\vcv||^2$.
\end{exercise}

\begin{exercise}{\hypertarget{unitproj}{Magnitude of Projection onto Unit Vector}}
Let $\vcv=(v_1,v_2,v_3)$ and $\vcu=(u_1,u_2,u_3)$ be vectors in $\bbr^3$. Show that if $\vcu$ is a unit vector, $||\proj{\vcv}{\vcu}||=|\vcv\bullet\vcu|$.
\end{exercise}
\renewcommand\thesubsection{\thesection.\arabic{subsection}}